---
title: "2025_02_04_Giotto_installation"
format: html
---

# Background and context

Elena will soon have data from a 10X Xenium run on her tonsil samples. This notebook will install Giotto and its dependencies.

It roughly follows the documentation from [Giotto documentation: Installation](https://giottosuite.readthedocs.io/en/latest/subsections/datasets/installation_04122023.html)

Giotto Suite is installed via R but there are required Python modules which must be installed in order for Giotto Suite to function properly. Below are the instructions for both the installation of the Giotto Suite package and required Python environment.

Requirements:

-   R (\>= 4.0.1)

-   Python (\>= 3.6)

-   Windows, MacOS or Linux specific installation tools for Posit (previously RStudio)

# Installation

## Installing Giotto Suite

```{r}
# Necessary for installation from R
if(!"devtools" %in% installed.packages()) {
  install.packages("devtools")
}

devtools::install_github("drieslab/Giotto@suite")
```

I ran into the following error:

> ERROR: dependencies ‘BiocParallel’, ‘BiocSingular’, ‘sparseMatrixStats’ are not available for package ‘Giotto’ \* removing ‘/Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library/Giotto’ Warning: installation of package ‘/var/folders/pc/74xx8xtj0796nbx_ns_m6kqw0000gp/T//RtmpFxyqqT/file13aa5aec1aa5/Giotto_4.2.0.tar.gz’ had non-zero exit status

I have a new laptop so I have not installed many packages on it yet. Let's try installing BiocParallel’, ‘BiocSingular’, ‘sparseMatrixStats’

```{r}
BiocManager::install(c("BiocParallel", "BiocSingular", "sparseMatrixStats"))
```

After this completed, I ran the devtools install command again and it worked with no errors.

## Install the Giotto python environment

```{r}
library(Giotto)
installGiottoEnvironment()
```
